---
title: "R for Biology Data Science Workshop Series"
subtitle: "Faculty of Natural and Agricultural Sciences, University of Pretoria"
author: "Joey Hulbert"
date: "April 23 - May 30, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE, warning=FALSE)
```

### Welcome! 

The purpose of this page is to provide resources for the *R for Biology Data Science* Workshop Series for postgrad students and postdocs in the [Faculty of Natural and Agricultural Sciences](https://www.up.ac.za/faculty-of-natural-agricultural-sciences) at the [University of Pretoria](https://www.up.ac.za/) in South Africa. 

* Workshop Series
    + April 23, 2019. **(Please note this session only starts at 4pm)**
        + [Session 1 - Introduction to RStudio](http://jmhulbert.github.io/r/bio/s1).
            + Learn the basics of coding, importing data, installing packages, and managing projects.
    + April 30, 2019. 
        + [Session 2](http://jmhulbert.github.io/r/bio/s2) - Introduction to data visualization with ggplot2 **Part 1**.
            + Create plots (histograms, boxplots, barplots), faceting, adjustments, etc.
    + May 7, 2019. 
        + Session 3 - Introduction to data visualization with ggplot2 **Part 2**.
            + Customizing plot aesthetics (colors, labels, italics, etc.), saving and exporting.
    + May 14, 2019.
        + **No session offered**
    + May 21, 2019. 
        + Session 4 - Introduction to data transformation with dplyr **Part 1**.
            + Filter data, summarize data (e.g. easily calculate means for groups), add variables. 
    + May 28, 2019. 
        + Session 5 - Introduction to data transformation with dplyr **Part 2**.
            + Merge datasets, combine tables, introduction to stringr, introduction to R markdown. 


### Pre-workshop instructions           

Please check out the [pre-workshop instructions](http://jmhulbert.github.io/r/bio/pre) prior to attending the first workshop (Link coming soon).


## Details {.tabset .tabset-fade}

### Schedule

**The first session will only start at 4pm**, all other sessions will start at 3pm. 

Please note that there is no session scheduled for May 14th. 

##### Google calendar

Here is [a link](https://calendar.google.com/calendar?cid=dXAuYWMuemFfbzR2ZTdmMWk1bzU2dG00bTlkMzNvb20ybGNAZ3JvdXAuY2FsZW5kYXIuZ29vZ2xlLmNvbQ) to the Google calendar for the workshop series. You can add the calendar to your diary. 

https://calendar.google.com/calendar?cid=dXAuYWMuemFfbzR2ZTdmMWk1bzU2dG00bTlkMzNvb20ybGNAZ3JvdXAuY2FsZW5kYXIuZ29vZ2xlLmNvbQ

<iframe src="https://calendar.google.com/calendar/embed?showPrint=0&amp;showTabs=0&amp;showCalendars=0&amp;showTz=0&amp;height=300&amp;wkst=2&amp;bgcolor=%23FFFFFF&amp;src=up.ac.za_o4ve7f1i5o56tm4m9d33oom2lc%40group.calendar.google.com&amp;color=%230D7813&amp;ctz=Africa%2FJohannesburg" style="border-width:0" width="400" height="300" frameborder="0" scrolling="no"></iframe>

### Location

All sessions will be offered in the auditorium of the Forestry and Agricultural Biotechnology Institute (FABI).

##### Directions 
The audiotorium is on the bottom floor of the FABI 2 building, but you will need to enter from the FABI 1 building. Head left from the front door and continue to the left (two additional left turns) until you reach the auditorium from the front door. 

<iframe src="https://www.google.com/maps/embed?pb=!1m18!1m12!1m3!1d3593.4428579872397!2d28.233133715020852!3d-25.75593198363941!2m3!1f0!2f0!3f0!3m2!1i1024!2i768!4f13.1!3m3!1m2!1s0x1e9561bb6399433b%3A0x53cd0dfb3f8fa93b!2sForestry+and+Agricultural+Biotechnology+Institute!5e0!3m2!1sen!2sza!4v1555584912565!5m2!1sen!2sza" width="400" height="300" frameborder="0" style="border:0" allowfullscreen></iframe>

##

## Participant Information Breakdown {.tabset .tabset-fade}

```{r include=FALSE}
library(tidyverse)
rsvp <- read.csv("./data/rsvp data.csv")
names(rsvp)
```

### Department

```{r}
ggplot(rsvp,aes(Department,fill=Department))+geom_histogram(stat="count")+coord_flip()+guides(fill="none")+theme_bw()
```

### Academic Standing

```{r}
ggplot(rsvp,aes(Academic.Standing,fill=Academic.Standing))+geom_histogram(stat="count")+coord_flip()+guides(fill="none")+theme_bw()
```

### Year

Year in academic position

```{r}
ggplot(rsvp,aes(Year.in.current.position.,fill=Year.in.current.position.))+geom_histogram(stat="count")+coord_flip()+guides(fill="none")+theme_bw()
```

### Data availability

Do you have your own data? 

```{r}
ggplot(rsvp,aes(Do.you.have.your.own.data.,fill=Do.you.have.your.own.data.))+geom_histogram(stat="count")+coord_flip()+guides(fill="none")+theme_bw()
```

### Computer Availability

Do you have a computer you can bring to the sessions?

```{r}
ggplot(rsvp,aes(Do.you.have.a.computer.you.can.bring.to.the.sessions.,fill=Do.you.have.a.computer.you.can.bring.to.the.sessions.))+geom_histogram(stat="count")+coord_flip()+guides(fill="none")+theme_bw()
```

### R Familiarity

Have you used R before?

```{r}
ggplot(rsvp,aes(Have.you.used.R.before.,fill=Have.you.used.R.before.))+geom_histogram(stat="count")+coord_flip()+guides(fill="none")+theme_bw()
```

If so, how often? (1=never, 10=everday)

```{r}
ggplot(rsvp,aes(If.so..how.often..,fill=If.so..how.often..))+geom_histogram(stat="count")+guides(fill="none")+theme_bw()
```

## 

