---
title: "Joseph (Joey) Michael Hulbert"
date: "CV Updated: `r format(Sys.time(), '%d %B, %Y')`"
subtitle: "WSU Urban Forest Health Lab,\n\nPuyallup Research and Extension Center,
  \n\nWashington State University, \n\n2606 West Pioneer, Puyallup, WA, 98371-4998
  \n\nhttps://jmhulbert.github.io \n\n1-541-908-5129 \n\nhulbe@wsu.edu\n\n[Curriculum Vitae PDF](https://jmhulbert.github.io/hulbert-cv.pdf)"
output:
  pdf_document:
    pandoc_args: !expr c(paste0("--lua-filter=", normalizePath("italicize_pathogens.lua")))
    toc: yes
    toc_depth: 5
  html_document:
    pandoc_args: !expr c(paste0("--lua-filter=", normalizePath("italicize_pathogens.lua")))
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
library(knitr)
library(kableExtra)
library(googlesheets4)
library(tidyverse)
library(scales)
```


```{r}
## Use this if you just want to read and not write to google drive

#gs4_deauth()
#df <- read_sheet("https://docs.google.com/spreadsheets/d/XXXXXXXXXXXX")
```


```{r load_google_sheets, warning=FALSE, message=FALSE,cache=TRUE}
gs4_deauth()
Initiatives <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Initiatives",.name_repair = ~ gsub(" ", ".", .x))
Digital.Media <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Digital Media Production",.name_repair = ~ gsub(" ", ".", .x))
Pop.Writing <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Pop Writing",.name_repair = ~ gsub(" ", ".", .x))
Media.Coverage <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Media Coverage",.name_repair = ~ gsub(" ", ".", .x))
Teaching.Experience <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Teaching Experience",.name_repair = ~ gsub(" ", ".", .x))
Science.Service <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Science Community Service",.name_repair = ~ gsub(" ", ".", .x))
Engagement <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Engagement",.name_repair = ~ gsub(" ", ".", .x))
Mentorship<- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Mentorship and Supervision",.name_repair = ~ gsub(" ", ".", .x))
Funding <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Funding",.name_repair = ~ gsub(" ", ".", .x))
Presentations <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Presentations",.name_repair = ~ gsub(" ", ".", .x))
Publications <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Publications",.name_repair = ~ gsub(" ", ".", .x))
Professional.Development <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Professional Development",.name_repair = ~ gsub(" ", ".", .x))
Awards <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Awards",.name_repair = ~ gsub(" ", ".", .x))
Community.Service <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Community Service",.name_repair = ~ gsub(" ", ".", .x))
Leadership <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Academic Leadership",.name_repair = ~ gsub(" ", ".", .x))
Mentorship.Outcomes <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Mentorship Outcomes",.name_repair = ~ gsub(" ", ".", .x))
Research.Positions <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Research Positions",.name_repair = ~ gsub(" ", ".", .x))
Memberships <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Memberships",.name_repair = ~ gsub(" ", ".", .x))
Event.Org <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Event Organization",.name_repair = ~ gsub(" ", ".", .x))
References <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="References",.name_repair = ~ gsub(" ", ".", .x))
Internship.Program <- read_sheet("https://docs.google.com/spreadsheets/d/1InNjBSkJl9rUwBQomq3cFxTm4OE39gYTmCC3XlK-2Fw/edit?gid=1669197812#gid=1669197812",sheet="Internship Program",.name_repair = ~ gsub(" ", ".", .x))
```



```{r, include=FALSE}
#excel_sheets("./cv/data/hulbert-cv.xlsx")
#Initiatives <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Initiatives")
#Digital.Media <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Digital Media Production")
#Pop.Writing <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Pop Writing")
#Media.Coverage <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Media Coverage")
#Teaching.Experience <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Teaching Experience")
#Science.Service <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Science Community Service")
#Engagement <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Engagement")
#Mentorship<- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Mentorship and Supervision")
#Funding <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Funding")
#Presentations <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Presentations")
#Publications <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Publications")
#Professional.Development <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Professional Development")
#Awards <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Awards")
#Community.Service <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Community Service")
#Leadership <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Academic Leadership")
#Mentorship.Outcomes <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Mentorship Outcomes")
#Research.Positions <- read_excel("./cv/data/hulbert-cv.xlsx",sheet="Research Positions")
```

## Academic Appointments

+ 2024-2026. **Research Assistant Professor (Career Track)**, Department of Plant Pathology, Washington State University.
+ 2022-2024. **Postdoctoral Research Associate**, WSU Puyallup Research and Extension Center, Washington State University.
+ 2020-2022. **USDA NIFA Postdoctoral Fellow**, Ornamental Plant Pathology, WSU Puyallup Research and Extension Center, Washington State University.

## Education

+ 2015-2020. **PhD** Plant Pathology, Forestry and Agricultural Biotechnology Institute, Dept. of Plant & Soil Science, University of Pretoria, South Africa. Dissertation: Phytophthora diversity in the Cape Floristic Region. Primary Supervisor: Mike Wingfield. Co-Supervisors: Treena Burgess, Francois Roets.
+ 2016-2019. **Affiliate Research Student**, Dept. of Conservation Ecology & Entomology, Stellenbosch University, South Africa.
+ 2012-2014. **Dual MSc** Dept. of Botany & Plant Pathology, Dept. of Wood Science & Engineering, Oregon State University. Thesis: Phytosanitation of Douglas-fir (Pseudotsuga menziesii) logs to mitigate the risk of disseminating Phytophthora ramorum within the global trade of timber. Advisers: Everett Hansen & Jeff Morrell.
+ 2006-2010. **BSc** Dept. of Natural Resource Sciences, Washington State University. Undergraduate Thesis: Influence of climatic variables on forest growth in the Sierra Nevada Mountains.

## Initiatives

`r paste0("- ", Initiatives$Year,",", " ", Initiatives$Initiative,ifelse(is.na(Initiatives$url) | Initiatives$url == "","", paste0("  \n  [", Initiatives$url, "](", Initiatives$url, ")")),collapse = "\n")`


## Community Service

`r paste0("- ", Community.Service$Year, "."," ", Community.Service$Service, collapse = "\n")`

## Publications

### Peer Reviewed

`r paste0("- ", Publications$Year[Publications$Type=="Peer-Reviewed Manuscript"],"."," ", Publications$Publication[Publications$Type=="Peer-Reviewed Manuscript"], collapse = "\n")`

### Not Peer Reviewed

`r paste0("- ", Publications$Year[Publications$Type=="Not Peer-Reviewed Manuscript"],"."," ", Publications$Publication[Publications$Type=="Not Peer-Reviewed Manuscript"], collapse = "\n")`

### Book Chapters

`r paste0("- ", Publications$Year[Publications$Type=="Book Chapter"],"."," ", Publications$Publication[Publications$Type=="Book Chapter"], collapse = "\n")`

### Extension Publications

`r paste0("- ", Publications$Year[Publications$Type=="Extension Fact Sheet"],"."," ", Publications$Publication[Publications$Type=="Extension Fact Sheet"], collapse = "\n")`

### Community Scientist Field Guides

`r paste0("- ", Publications$Year[Publications$Type=="Community Science Field Guide"],"."," ", Publications$Publication[Publications$Type=="Community Science Field Guide"], collapse = "\n")`

### ArcGIS Story Maps

`r paste0("- ", Publications$Year[Publications$Type=="ArcGIS Story Map"],"."," ", Publications$Publication[Publications$Type=="ArcGIS Story Map"], collapse = "\n")`

### Progress Reports

`r paste0("- ", Publications$Year[Publications$Type=="Program Progress Report"],"."," ", Publications$Publication[Publications$Type=="Program Progress Report"], collapse = "\n")`

## Funding and Support

### Total Funding Requests

```{r, message=FALSE}

funding.summary <- Funding %>% filter(Funding.Type!="In Kind Donations") %>%
  group_by(Funding.Type) %>%
  summarize(
    n = n(),
    n.Funded=sum(Funding.Status == "Funded",na.rm=TRUE),
    Total = sum(Amount, na.rm = TRUE),
    Funded = sum(if_else(Funding.Status == "Funded", Amount, 0), na.rm = TRUE),
    Pending = sum(if_else(Funding.Status == "Pending", Amount, 0), na.rm = TRUE),
    Funded.as.PI = sum(if_else(Funding.Status=="Funded" & Role == "PI", Amount, 0), na.rm = TRUE),
    .groups = "drop"
  ) %>% arrange(desc(Total)) %>%
 bind_rows(
    Funding %>%
      summarize(
        Funding.Type = "Total",
        n = n(),
        n.Funded=sum(Funding.Status == "Funded",na.rm=TRUE),
        Total = sum(Amount, na.rm = TRUE),
        Funded = sum(if_else(Funding.Status == "Funded", Amount, 0), na.rm = TRUE),
        Pending = sum(if_else(Funding.Status == "Pending", Amount, 0), na.rm = TRUE),
        Funded.as.PI = sum(if_else(Funding.Status=="Funded" & Role == "PI", Amount, 0), na.rm = TRUE),
      )
  ) %>% mutate(across(c(Total, Funded, Pending, Funded.as.PI), dollar))

kable(
  funding.summary,
  align=c("l"),
  format = "latex",   # or "html"
  booktabs = TRUE,
 ) %>%
  kable_styling(
    latex_options = c("hold_position"),
    full_width = FALSE,
    position = "left"
  )
```

### Total Funding Awarded

```{r, message=FALSE}

funded <- Funding %>% filter(Funding.Status=="Funded") %>% filter(Funding.Type!="In Kind Donations") %>% 
  group_by(Year.Awarded) %>%
  summarize(
    n = n(),
    Total = sum(Amount, na.rm = TRUE),
    PI = sum(if_else(Role == "PI", Amount, 0), na.rm = TRUE),
    Co.PI = sum(if_else(Role == "Co-PI", Amount, 0), na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  arrange(desc(Year.Awarded)) %>% 
  mutate(Year.Awarded = as.character(Year.Awarded)) %>%
bind_rows(
    summarize(
      .,
      Year.Awarded = "Total",
      across(where(is.numeric), sum)
    )
  ) %>% 
  mutate(across(c(Total,PI,Co.PI), dollar))

kable(
  funded,
  align=c("l"),
  format = "latex",   # or "html"
  booktabs = TRUE,
 ) %>%
  kable_styling(
    latex_options = c("hold_position"),
    full_width = FALSE,
    position = "left"
  )
```

### Active Funded Projects
n= `r nrow(Funding[Funding$Project.Status=="Active",])`

```{r}
Funded <- Funding %>% filter(Project.Status=="Active") %>% mutate(Amount=dollar(Amount)) %>% arrange(desc(Year.Awarded))
```


`r paste0("- ", Funded$Amount,"."," ", Funded$Role,"."," ", Funded$Year.Awarded,"."," ", Funded$Source,"."," ",Funded$Title,  collapse = "\n")`

### Sources of Funding Awarded

```{r, message=FALSE}

funded.source <- Funding %>% filter(Funding.Status=="Funded") %>% filter(Funding.Type!="In Kind Donations") %>%
  group_by(Source) %>%
  summarize(
    n = n(),
    Total = sum(Amount, na.rm = TRUE),
    PI = sum(if_else(Role == "PI", Amount, 0), na.rm = TRUE),
    Co.PI = sum(if_else(Role == "Co-PI", Amount, 0), na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  arrange(desc(Total)) %>% 
bind_rows(
    summarize(
      .,
      Source = "Total",
      across(where(is.numeric), sum)
    )
  ) %>% 
  mutate(across(c(Total,PI,Co.PI), dollar))

kable(
  funded.source,
  align=c("l"),
  format = "latex",   # or "html"
  booktabs = TRUE,
 ) %>%
  kable_styling(
    latex_options = c("hold_position"),
    full_width = FALSE,
    position = "left"
  )
```

### In Kind Donations

```{r}
In.Kind <- Funding %>% filter(Funding.Type=="In Kind Donations") %>% mutate(Amount=dollar(Amount)) %>% arrange(desc(Year.Awarded))
```


`r paste0("- ", In.Kind$Amount,"."," ", In.Kind$Source,"."," ",In.Kind$Title,".",  collapse = "\n")`


## Academic Honors and Awards

`r paste0("- ", Awards$Year, "."," ", Awards$Award, collapse = "\n")`

## Academic Leadership

`r paste0("- ", Leadership$Year, "."," ", Leadership$Role, collapse = "\n")`


## Science Community Service

`r paste0("- ", Science.Service$Year, "."," ", Science.Service$Service, collapse = "\n")`

## Memberships

`r paste0("- ", Memberships$Year, "."," ", Memberships$Membership, collapse = "\n")`

## Teaching Experience


`r paste0("- ", Teaching.Experience$Year, "."," ", Teaching.Experience$Experience, "."," ",  Teaching.Experience$Location, collapse = "\n")`

### Internship Program

`r paste0("- ",Internship.Program$Year, ". ","Week ", Internship.Program$Week, ". ",Internship.Program$Topic, ". ", ifelse(!is.na(Internship.Program$Guest),paste0("Featuring: ",Internship.Program$Guest),""),".", collapse = "\n")`



## Event Organization
n= `r nrow(Event.Org)`

`r paste0("- ",  Event.Org$Year, ". ",  Event.Org$Title, ".",  if_else(!is.na(Event.Org$Number.of.Participants) & Event.Org$Number.of.Participants > 0, paste0(" ", Event.Org$Number.of.Participants, " Participants."),""  ),  collapse = "\n")`

## Presentations

### Summary

```{r, message=FALSE,warning=FALSE}
n_type_by_year <- Presentations %>%
  count(Presentation.Type.Simplified, Year, name = "n") %>%
  pivot_wider(
    names_from  = Year,     # years become columns
    values_from = n,
    values_fill = 0
  ) %>%
  arrange(Presentation.Type.Simplified)

year_cols <- sort(
  as.integer(setdiff(names(n_type_by_year), "Presentation.Type")),
  decreasing = TRUE
)


n_type_by_year <- n_type_by_year %>%
  select(Presentation.Type.Simplified, all_of(as.character(year_cols))) %>%
  arrange(Presentation.Type.Simplified)

total_row <- n_type_by_year %>%
  summarise(
    Presentation.Type.Simplified = "Total in Year",
    across(where(is.numeric), sum)
  )

n_type_by_year <- bind_rows(n_type_by_year, total_row)


kable(
  n_type_by_year,
  col.names = c(
    "Presentation Type",
    names(n_type_by_year)[-1]),
  align = c("l", rep("c", ncol(n_type_by_year) - 1)),
  format = "latex",
  booktabs = TRUE,
  escape = FALSE
) %>%
  kable_styling(
    latex_options = "hold_position",
    full_width = FALSE,
    position = "left"
  ) %>%
  column_spec(1, width = "1.1in") %>%                 # Presentation.Type column
  column_spec(2:ncol(n_type_by_year), width = "0.18in")
```


### Conference & Symposia Presentations
n= `r nrow(Presentations  %>%  filter(Presentation.Type.Simplified %in% c("Conference", "Symposia")))`

`r paste0("- ", Presentations %>%  filter(Presentation.Type.Simplified %in% c("Conference", "Symposia")) %>%  arrange(desc(Year)) %>%   pull(Year), ". ", Presentations %>%   filter(Presentation.Type.Simplified %in% c("Conference", "Symposia")) %>%   arrange(desc(Year)) %>%   pull(Title),collapse = "\n")`

### 2025 Presentations
n= `r nrow(Presentations %>% filter(Year==2025))`


`r paste0("- ", Presentations %>%  filter(Year==2025) %>%  pull(Year), ". ", Presentations %>%  filter(Year==2025) %>% pull(Title),collapse = "\n")`

## Engagement

### Summary

```{r, message=FALSE, warning=FALSE}
n_activity_by_year <- Engagement %>%   # <- your table with Engagement Activity / Year / Date / Event
  count(Engagement.Activity, Year, name = "n") %>%
  pivot_wider(
    names_from  = Year,      # years become columns
    values_from = n,
    values_fill = 0
  )

# reorder year columns descending (left → right)
year_cols <- sort(
  as.integer(setdiff(names(n_activity_by_year), "Engagement Activity")),
  decreasing = TRUE
)

n_activity_by_year <- n_activity_by_year %>%
  select(Engagement.Activity, all_of(as.character(year_cols))) %>%
  arrange(Engagement.Activity)

# add Total row
total_row <- n_activity_by_year %>%
  summarise(
    Engagement.Activity = "Total in Year",
    across(where(is.numeric), sum)
  )

n_activity_by_year <- bind_rows(n_activity_by_year, total_row)

kable(
  n_activity_by_year,
  col.names = c("Engagement Activity", names(n_activity_by_year)[-1]),
  align = c("l", rep("c", ncol(n_activity_by_year) - 1)),
  format = "latex",
  booktabs = TRUE,
  escape = FALSE
) %>%
  kable_styling(
    latex_options = "hold_position",
    full_width = FALSE,
    position = "left"
  ) %>%
  column_spec(1, width = "1.6in") %>%
  column_spec(2:ncol(n_activity_by_year), width = "0.18in")
```


### 2025 Engagement Activities

#### Youth Engagement
n= `r nrow(Engagement %>%  filter(Year==2025 & Engagement.Activity=="Youth Engagement"))`

`r paste0("- ", Engagement %>%  filter(Year==2025 & Engagement.Activity=="Youth Engagement") %>%  pull(Year), ". ", Engagement %>%  filter(Year==2025  & Engagement.Activity=="Youth Engagement") %>% pull(Event),".", collapse = "\n")`

#### Prmomotional Table Events
n= `r nrow(Engagement %>%  filter(Year==2025 & Engagement.Activity=="Promotional Table Events"))`

`r paste0("- ", Engagement %>%  filter(Year==2025 & Engagement.Activity=="Promotional Table Events") %>%  pull(Year), ". ", Engagement %>%  filter(Year==2025  & Engagement.Activity=="Promotional Table Events") %>% pull(Event),".", collapse = "\n")`

#### WSU Puyallup REC Tours
n= `r nrow(Engagement %>%  filter(Year==2025 & Engagement.Activity=="WSU Puyallup REC Tours"))`

`r paste0("- ", Engagement %>%  filter(Year==2025 & Engagement.Activity=="WSU Puyallup REC Tours") %>%  pull(Year), ". ", Engagement %>%  filter(Year==2025  & Engagement.Activity=="WSU Puyallup REC Tours") %>% pull(Event),".", collapse = "\n")`

#### Training Events
n= `r nrow(Engagement %>%  filter(Year==2025 & Engagement.Activity=="Training Event"))`


`r paste0("- ", Engagement %>%  filter(Year==2025 & Engagement.Activity=="Training Event") %>%  pull(Year), ". ", Engagement %>%  filter(Year==2025  & Engagement.Activity=="Training Event") %>% pull(Event),".", collapse = "\n")`

## Mentorship and Supervision

`r paste0("- ", Mentorship %>% arrange(desc(Year)) %>%   pull(Year), "."," ", Mentorship$Individual, " ", Mentorship$Type.of.Mentorship,".", collapse = "\n")`

### Mentorship Outcomes

`r paste0("- ", Mentorship.Outcomes %>% arrange(desc(Year)) %>%   pull(Year), "."," ", Mentorship.Outcomes$Outcome,".", collapse = "\n")`

## Professional Development

`r paste0("- ", Professional.Development %>% arrange(desc(Year)) %>%   pull(Year), "."," ", Professional.Development$Training, collapse = "\n")`


## Digital Media Production

`r paste0("- ", Digital.Media$Year, "."," ", Digital.Media$Type,"."," ", Digital.Media$Content, collapse = "\n")`


## Selected Popular Writing

`r paste0("- ", Pop.Writing$Year, "."," ", Pop.Writing$Authors,"."," ", Pop.Writing$Title,  collapse = "\n")`


## Selected Media Coverage


`r paste0("- ", Media.Coverage$Year, "."," ", Media.Coverage$Coverage.Type,"."," ", Media.Coverage$Source, collapse = "\n")`


## Research Positions

`r paste0("- ", Research.Positions$Year, "."," ", Research.Positions$Position, collapse = "\n")`

## References

`r paste0("- ", References$Year, "."," ", References$Individual, "."," ", References$Role, collapse = "\n")`