---
title: "Session 1 - Introduction to RStudio"
author: "Joey Hulbert"
date: "February 6, 2019"
output: 
  html_document:
    toc: TRUE 
    toc_depth: 5
    toc_float:
      collapsed: TRUE
---
<script>
    $(document).ready(function() {
      $items = $('div#TOC li');
      $items.each(function(idx) {
        num_ul = $(this).parentsUntil('#TOC').length;
        $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
      });
    });
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Introduction
 
The purpose of this page is to provide the resources for the first session of our RStudio and R::packages workshop series. This page is designed to compliment our workshop session and is not supposed to be a stand alone walkthrough, however, it may work that way. 

##Organize data appropriately


##Install R and RStudio

* Before we begin, we need to install [R](https://www.r-project.org/) and [RStudio](https://www.rstudio.com/) (RStudio requires R to be installed).
    + Download R: https://ftp.osuosl.org/pub/cran/
    + Download RStudio: https://www.rstudio.com/products/rstudio/download/#download

It is recommended you have the latest versions of R (3.5.2) and RStudio (1.1.463). 

To find your RStudio version go to the menu: RStudio >> About RStudio

Run the below code to find your R version. 

```{r}
getRversion()
```

or

```{r}
sessionInfo()
```



##RStudio interface

* Arrange Panes
    + Source
    + Console
    + Environment, Plots, Help, Viewer
    + History, Files, Connections, Packages

Appearance


##Packages

Install tidyverse


##Projects

Set working directories

Create R script

Importing data

name <- read.csv('./data/'FILENAME.csv')

Load Packages

tidyverse

R code basics

Make notes

Use outline

Reference column in data

ls() list everything in environment

rm() removes from environment

str()
names()
levels(data$column)

rename a column
x$y <- x$z

Change data from R

library(stringr)
arthropod$Cultivar <-str_replace_all(arthropod$Cultivar,"Jewel ","Jewel")
arthropod$Cultivar <-str_replace_all(arthropod$Cultivar,"Star ","Star")
arthropod$Field <-str_replace_all(arthropod$Field,"Open ","Open")



##Package Cheat Sheets


