---
title: "FABI Session 4 - Introduction to vegan"
subtitle: "Forestry and Agricultural Biotechnology Institute, University of Pretoria"
author: "Joey Hulbert"
date: "March 7, 2019"
output: html_document
---
<script>
    $(document).ready(function() {
      $items = $('div#TOC li');
      $items.each(function(idx) {
        num_ul = $(this).parentsUntil('#TOC').length;
        $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
      });
    });
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(knitr)
```

[FABI Workshop Series >> Home](http://jmhulbert.github.io/r)
[FABI Workshop Series >> Mock Data](http://jmhulbert.github.io/r/data)


# Introduction

Welcome to the guide for the fourth session of our workshop. The purpose of this page is to provide an introduction to the [vegan package](https://www.rdocumentation.org/packages/vegan/versions/2.4-2) and provide an example of a multivariate analysis. 

Then content of this session is available from Joey. Please contact him if you're interested in the content from this session. Thanks. 

