---
title: "FABI Pre-Workshop Series Instructions"
author: "Joey Hulbert"
date: "February 13, 2019"
output: html_document
---

<script>
    $(document).ready(function() {
      $items = $('div#TOC li');
      $items.each(function(idx) {
        num_ul = $(this).parentsUntil('#TOC').length;
        $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
      });
    });
</script>

[FABI Workshop Series >> Home](http://jmhulbert.github.io/r)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(knitr)
```


## Welcome
 
Thank you for RSVPing to attend the FABI Rstudio and R Package workshop. Below are a few instructions that will improve the experience if you can incorporate them prior to attending the workshop sessions.

## Data organization

We need to format our data so that R will read it correctly. 

Our data should be set up as a rectangle with observations (samples) in rows and variables (treatment groups, results, etc.) in columns. Below is an example:

```{r echo=FALSE}
df <- data.frame(Sample = 1:12,Trial = c('A','A','A','A','A','A','B','B','B','B','B','B'),Treatment = rep(c('Control','Control','Control','Treated','Treated','Treated'),2),Result = runif(12))
knitr::kable(df,align='c')
```

It is best to import your raw data and not bother calculating averages or anything in excel previously. We will use dplyr in [Session 3](http://jmhulbert.github.io/r/s3) to summarize data and calculate means based on groups or other explanatory variables, etc. 

Now that our data is ready for R, we need to install R and RStudio.

## Install R and RStudio
* Note that [RStudio](https://www.rstudio.com/) requires [R](https://www.r-project.org/) to be installed.
    + First download and install R: https://ftp.osuosl.org/pub/cran/
    + Then download and install RStudio: https://www.rstudio.com/products/rstudio/download/#download

It is best if we have the latest versions of R (3.5.2) and RStudio (1.1.463). 

To find your RStudio version go to the menu: *RStudio >> About RStudio*

Run the below code to find your R version. 

```{r eval=FALSE}
getRversion()
```

or

```{r eval=FALSE}
sessionInfo()
```
