---
title: "R for Biology Data Science - Session 2 - Introduction to ggplot2 part 1"
subtitle: "Faculty of Natural and Agricultural Sciences, University of Pretoria"
author: "Joey Hulbert"
date: "April 30, 2019"
output: 
  html_document:
    css: custom.css
    toc: TRUE 
    toc_depth: 5
    toc_float:
      collapsed: FALSE
---
<script>
    $(document).ready(function() {
      $items = $('div#TOC li');
      $items.each(function(idx) {
        num_ul = $(this).parentsUntil('#TOC').length;
        $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
      });
    });
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(knitr)
```

* Workshop Series
    + [Home](http://jmhulbert.github.io/r/bio/)
    + [Pre-session instructions](http://jmhulbert.github.io/r/bio/pre)
    + [Session 1](http://jmhulbert.github.io/r/bio/s1)
    + [Session 2](http://jmhulbert.github.io/r/bio/s2)
    + [Mock Data](http://jmhulbert.github.io/r/bio/data)
    
## Introduction

Welcome to the second session of our workshop series. The purpose of this page is to provide a guide through the subjects we will cover. In this session we will introduce a powerful, highly customizable, and widely used package for visulizing data, [ggplot2](https://ggplot2.tidyverse.org/).

## Pre-Session Instructions

You will need to have RStudio (and R) installed on your computer and have the 'tidyverse' package installed. More information about this can be found in the lesson for [Session 1](http://jmhulbert.github.io/r/bio/s1).

### Install Packages

Remember you can install packages using the below code.

```{r eval=FALSE}
install.packages("tidyverse")
```

It can take a few minutes for RStudio to unpack/install the packages. The [Tidyverse](https://www.tidyverse.org/) package is a collection (package) of packages. It includes several individual packages that work fairly well together. 


More information coming soon.
