---
title: "R for Biology Data Science - Session 5 - Introduction to dplyr part 2"
subtitle: "Faculty of Natural and Agricultural Sciences, University of Pretoria"
author: "Joey Hulbert"
date: "5/21/2019"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
---

<script>
    $(document).ready(function() {
      $items = $('div#TOC li');
      $items.each(function(idx) {
        num_ul = $(this).parentsUntil('#TOC').length;
        $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
      });
    });
</script>

* Workshop Series
    + [Home](http://jmhulbert.github.io/r/bio/)
    + [Pre-session instructions](http://jmhulbert.github.io/r/bio/pre)
    + [Session 1](http://jmhulbert.github.io/r/bio/s1)
    + [Session 2](http://jmhulbert.github.io/r/bio/s2)
    + [Session 3](http://jmhulbert.github.io/r/bio/s3)
    + [Session 4](http://jmhulbert.github.io/r/bio/s4)
    + [Mock Data](http://jmhulbert.github.io/r/bio/data)
    
```{r include=FALSE}
library(knitr)
```


## Introduction

Welcome to the fifth session of our workshop series. The purpose of this page is to serve as a guide for the material we will cover during the session. We will recap on methods to summarize and transform data using the [dplyr](https://dplyr.tidyverse.org/) package and have a quick introduction to R markdown. 

```{r include=FALSE}
library(tidyverse)
```

```{r eval=FALSE}
library(tidyverse)
```


### Cheatsheets

Remember there is a nice cheatsheet for dplyr available here: [dplyr cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf) - click to download.

More cheatsheets for other packages and updates can be found here: [https://www.rstudio.com/resources/cheatsheets/](https://www.rstudio.com/resources/cheatsheets/).

### Index our data
```{r}
data <- read.csv('Silver Tree Study.csv')
```


### Filter

```{r}
data.filtered <- data %>% filter(Photosynthesis<300)
```


```{r}
nrow(data)
nrow(data.filtered)
```

Now we can see 10 rows have been effectively filtered off.

### Mutate

```{r}
data.filtered.new.column <- data %>% mutate(new.column = Photosynthesis/Conductance)
```


```{r}
ncol(data.filtered)
ncol(data.filtered.new.column)
```

Now we can see there is an additional column

### Summarize

```{r}
Treatment.mean <- data.filtered %>% group_by(Treatment,Species) %>% summarize(n=n(), mean=mean(Photosynthesis),sd=sd(Photosynthesis),se = sd / sqrt(n),lowse = (mean-se),highse = (mean+se))
knitr::kable(Treatment.mean,align="c") ## note the kniter:: is telling r to look for the kable command in the knitter package. If you index(load) the knittr package at the beginning of your session, you can just write kable()
```


### Merge datasets

Lets calculate the proportion of plants in each treatment on each day measured. 

First we need to calculate the total number of plants on each day

```{r}
Total.plants.per.day <- data.filtered %>% group_by(Days.after.inoculation) %>% summarize(Total=n_distinct(Unique.Sample.Number))
```

Now lets calculate the number of plants in each treatment per day

```{r}
Plants.per.treatment.per.day <- data.filtered %>% group_by(Treatment,Days.after.inoculation) %>% summarize(Number.of.Plants=n_distinct(Unique.Sample.Number))
```

Now we can merge the summary tables based on days after inoculation

```{r} 
Plants.overall <- left_join(Plants.per.treatment.per.day,Total.plants.per.day,by="Days.after.inoculation") #join matching values from total.plants.per.day to plants.per.treatment.per.day)
```

Now we can caclulate the proportions

```{r}
Plants.overall <- Plants.overall %>% mutate(Proportion=Number.of.Plants/Total)
kable(Plants.overall,align="c")
```

hmm lets round the propotion values by adding the round() command to above code

```{r}
Plants.overall <- Plants.overall %>% mutate(Proportion=round(Number.of.Plants/Total,2))
kable(Plants.overall,align="c")
```

### Other handy packages in tidyverse

#### Stringr


"The [stringr](https://stringr.tidyverse.org/) package provides an easy to use toolkit for working with strings, i.e. character data, in R. This cheatsheet guides you through stringr’s functions for manipulating strings. The back page provides a concise reference to regular expresssions, a mini-language for describing, finding, and matching patterns in strings." - [RStudio](https://www.rstudio.com/resources/cheatsheets/). 

```{r}
data.with.new.column <- data.filtered %>% mutate(and.and.and = "one & two & three") #and.and.and is the name of the new column, which is just a 'string' of text copied in every row.
levels(data.with.new.column$and.and.and)

data.new.rows <-separate_rows(data.with.new.column, and.and.and, sep = "&")
levels(data.new.rows$and.and.and)
```



#### forcats


"Factors are R’s data structure for categorical data. The [forcats](https://forcats.tidyverse.org/) package makes it easy to work with factors. This cheatsheet reminds you how to make factors, reorder their levels, recode their values, and more" - [RStudio](https://www.rstudio.com/resources/cheatsheets/).


#### R Markdown

"R Markdown is an authoring format that makes it easy to write reusable reports with R. You combine your R code with narration written in markdown (an easy-to-write plain text format) and then export the results as an html, pdf, or Word file. You can even use R Markdown to build interactive documents and slideshows. " - [RStudio](https://www.rstudio.com/resources/cheatsheets/)

* Here are some critical links
    + [R Markdown cheatsheet](https://content.cdntwrk.com/files/aT05NjI5MDcmdj0xJmlzc3VlTmFtZT1ybWFya2Rvd24tMi0wLWNoZWF0LXNoZWV0JmNtZD1kJnNpZz0xYTkwYWM1ZjUzMjVhZjRkYTFmOWFmYmM1ZTcwZjZiMw%253D%253D)
    + [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/)
    
    
